---
title: "lab_07_group15.qmd"
format:
  html:
    embed-resources: true
editor: visual
---

Group 15:

-   José Dinis Coelho da Silva Ferreira, s212484

-   Maria Gabriela Frascella, s233113

-   Paula Gomez-Plana Rodriguez, s233165

-   Alicia Llorente Herrero, s233658

-   Ona Saulianskaite, s232958

## Background 

The dataset used in this project is called biopsy. This dataset is from the University of Wisconsin Hospitals, and it contains biopsies of breast cancer of 699 patients. Each observation has nine attributes in a scale of 1 to 10 and a tenth attribute with the outcome (benign/ malignant).

## Load packages

```{r}

library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)
```

## Load data

```{r}
biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

## Analysis

```{r}
pca_fit <- biopsy %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(scale = TRUE) # do PCA on scaled data
```

```{r}
pca_fit %>%
  augment(biopsy) %>% # add original dataset back in
  ggplot(aes(.fittedPC1, .fittedPC2, color = outcome)) + 
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(malignant = "#D55E00", benign = "#0072B2")
  ) +
  theme_half_open(12) + background_grid()
```
