---
title: "lab_07_group15.qmd"
format:
  html:
    embed-resources: true
editor: visual
---

Group 15:

-   José Dinis Coelho da Silva Ferreira, s212484

-   Maria Gabriela Frascella, s233113

-   Paula Gomez-Plana Rodriguez, s233165

-   Alicia Llorente Herrero, s233658

-   Ona Saulianskaite, s232958

## Overview of the dataset used

In this report, we will analyze a breast cancer biopsy dataset from the University of Wisconsin Hospitals, Madison, collected by Dr. William H. Wolberg. The dataset consists of biopsy data from 699 patients, each with nine attributes scored on a scale from 1 to 10. These attributes describe various characteristics of breast tumors, and each observation is classified as either malignant or benign, providing a valuable dataset for assessing and predicting tumor types.

## Loading libraries
```{r}
library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)
library(corrplot)
```

## Loading data
```{r}
biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

## Analysis

### Correlation analysis of different attributes
```{r}
# Corelation matrix without nominal attribute: "outcome"
biopsy |>
  select(-outcome) |>
  cor(method = "pearson") |>
  corrplot(method = "color", type = "full", tl.cex = 0.8, addCoef.col = "black", 
           number.cex = 0.7, number.digits = 2)
```

```{r}
pca_fit <- biopsy %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(scale = TRUE) # do PCA on scaled data
```

```{r}
pca_fit %>%
  augment(biopsy) %>% # add original dataset back in
  ggplot(aes(.fittedPC1, .fittedPC2, color = outcome)) + 
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(malignant = "#D55E00", benign = "#0072B2")
  ) +
  theme_half_open(12) + background_grid()
```
